---
title: "Example 1: Presenting Plots, Regular text and Sections"
author: Spencer Lourens, Ph.Daaaaaaaay
output: pdf_document
---

\section{Section 1} This is our first paragraph example in R Markdown. The file 1-example.Rmd can be opened in RStudio, and then "knitted" using alt + shift + K (Windows) or command + shift + K (MacOS). The following blocks show random generation of normal random variables and a histogram plot of this data

# This is a "Header 1"
## This is a "Header 2"
### This is a "Header 3"

## See [RMarkdown Basics](http://rmarkdown.rstudio.com/authoring_basics.html) for more basics in RMarkdown

## Inline R code: `r rnorm(1, 0,1)` was generated from a N(0,1) distribution. 

## Normal RVs histogram below: generated with an ``R code chunk"
```{r echo=FALSE}
## Notice the echo=FALSE above within the { }... this tells RMarkdown not to show the code but only the result.
normRVs <- rnorm(2000, 0, 1)
hist(normRVs, breaks = 20)
```

## How about a plot of exponential RVs? Notice that the code below is included in the document. This is because we do not include echo=FALSE within the { }s for this R code block

```{r}
expRVs <- rexp(2000, rate = 5)
hist(expRVs, breaks = 20)
```

If we don't want the text bolded, we can just type in the Rmd file as we would in usual LaTeX. Below we illustrate a R code chunk with the option ``eval = FALSE", which is very useful when you want to show the code you are using, but not evaluate the code.

```{r eval = FALSE}
expRVs <- rexp(2000, rate = 5)
hist(expRVs, breaks = 20)
```

Notice this did not produce another histogram plot, as opposed to the R code chunk above it.



