---
title: "R Data Wrangling"
author: "Spencer Lourens"
date: '`r format(Sys.Date(), "%B %d %Y")`'
output:
  beamer_presentation:
  theme: "Darmstadt"
  slide_level: 3
  includes:
    in_header: ~/header.txt
---

```{r library_inlusions, include = FALSE}
library(dplyr)
library(tidyr)
library(EDAWR)
library(nycflights13)
```

## Outline of Lecture
- Data Wrangling
- The dplyr package
    - Selecting, filtering observations
    - Calculating new variables
    - Summarizing the data
- The tidyr package
    - Converting wide to long
    - Converting long to wide


## Data Wrangling! (Using dplyr, tidyr)
- <b>Data wrangling</b>: The practical task of transforming the <em>layout</em>, <em>substance</em>, and <em>display</em> of your data - Garrett Grolemund, RStudio


## Data Wrangling! (Using dplyr, tidyr)
- <b>Data wrangling</b>: The practical task of transforming the <em>layout</em>, <em>substance</em>, and <em>display</em> of your data - Garrett Grolemund, RStudio
- <b>Data wrangling</b> can be broken into three basic domains: data manipulation, data tidying (for analysis/visualization), and data visualization


## Data Wrangling! (Using dplyr, tidyr)
- <b>Data wrangling</b>: The practical task of transforming the <em>layout</em>, <em>substance</em>, and <em>display</em> of your data - Garrett Grolemund, RStudio
- <b>Data wrangling</b> can be broken into three basic domains: data manipulation, data tidying (for analysis/visualization), and data visualization
    - <em>Data manipulation</em> - Change the variables, their values, or the sample units used for analysis (aggregation?)


## Data Wrangling! (Using dplyr, tidyr)
- <b>Data wrangling</b>: The practical task of transforming the <em>layout</em>, <em>substance</em>, and <em>display</em> of your data - Garrett Grolemund, RStudio
- <b>Data wrangling</b> can be broken into three basic domains: data manipulation, data tidying (for analysis/visualization), and data visualization
    - <em>Data manipulation</em> - Change the variables, their values, or the sample units used for analysis (aggregation?)
    - <em>Data tidying</em> - Change the LAYOUT of the data - wide, long, etc., most of the time for software use

## Data Wrangling! (Using dplyr, tidyr)
- <b>Data wrangling</b>: The practical task of transforming the <em>layout</em>, <em>substance</em>, and <em>display</em> of your data - Garrett Grolemund, RStudio
- <b>Data wrangling</b> can be broken into three basic domains: data manipulation, data tidying (for analysis/visualization), and data visualization
    - <em>Data manipulation</em> - Change the variables, their values, or the sample units used for analysis (aggregation?)
    - <em>Data tidying</em> - Change the LAYOUT of the data - wide, long, etc., most of the time for software use
    - <em>Data visualization</em> - Make use of our visual way of thinking by representing data in a visual format

## Preliminaries
- The dplyr package creates another structure, called a tbl (table, or as Hadley says, tibble)


## Preliminaries
- The dplyr package creates another structure, called a tbl (table, or as Hadley says, tibble)
-  A tbl is a dataframe 

## Preliminaries
- The dplyr package creates another structure, called a tbl (table, or as Hadley says, tibble)
- A tbl is a dataframe 
    - that works seemlessly with dplyr functions

## Preliminaries
- The dplyr package creates another structure, called a tbl (table, or as Hadley says, tibble)
- A tbl is a dataframe 
    - that works seemlessly with dplyr functions
    - that optimizes display when printed without printing too much and taking up unnecessary cache memory (slows down your R session)

## Preliminaries
- The dplyr package creates another structure, called a tbl (table, or as Hadley says, tibble)
- A tbl is a dataframe 
    - that works seemlessly with dplyr functions
    - that optimizes display when printed without printing too much and taking up unnecessary cache memory (slows down your R session)
- The pipe (%>%) operator

## Preliminaries
- The dplyr package creates another structure, called a tbl (table, or as Hadley says, tibble)
- A tbl is a dataframe 
    - that works seemlessly with dplyr functions
    - that optimizes display when printed without printing too much and taking up unnecessary cache memory (slows down your R session)
- The pipe (%>%) operator
    - allows more concise, perhaps narrative-looking production of code

## Piping example
- Consider the two approaches below
```{r}
x <- c(1,3,4,1,4,10,19,2.5,1)
x %>% mean(na.rm = T)
mean(x)
```
- Same result!


## Piping example (Cont'd)
- Piping really shines when you have many nested calls as in toy example below

## Piping example (Cont'd)
- Piping really shines when you have many nested calls as in toy example below
- The following are equivalent:
    - reportResult(groupResult(calculateResult(df))
    - df %>% calculateResult() %>% groupResult() %>% reportResult()

## Piping example (Cont'd)
- Piping really shines when you have many nested calls as in toy example below
- The following are equivalent:
    - reportResult(groupResult(calculateResult(df))
    - df %>% calculateResult() %>% groupResult() %>% reportResult()
- Notice that the second line with piping is more natural to read from left to right, similar to how we read English

## Piping example (Cont'd)
- Piping really shines when you have many nested calls as in toy example below
- The following are equivalent:
    - reportResult(groupResult(calculateResult(df))
    - df %>% calculateResult() %>% groupResult() %>% reportResult()
- Notice that the second line with piping is more natural to read from left to right, similar to how we read English
- Very helpful in more complicated settings

## Example datasets
- It may be helpful to go back over these notes in your own R session - we'll test a lot of dplyr functionality on different datasets

## Example datasets
- It may be helpful to go back over these notes in your own R session - we'll test a lot of dplyr functionality on different datasets
- Functions from the dplyr package to demonstrate: select, filter, mutate, group_by, summarise ... 

## Example datasets
- It may be helpful to go back over these notes in your own R session - we'll test a lot of dplyr functionality on different datasets
- Functions from the dplyr package to demonstrate: select, filter, mutate, group_by, summarise ... 
- We will illustrate use of these functions on the flights data set from package <b>nycflights13</b>, previewed on the following slide, and the storms data set from package <b>EDAWR</b> (and other EDAWR datasets), on the subsequent slide.

## Dataset flights
- The actual dimension of flights is 336776 by 19
```{r echo = FALSE, results = 'asis'}
knitr::kable(flights[1:10, 1:7])
```

## Dataset storms
- The actual dimension of storms is 6 by 4
```{r echo = FALSE, results = 'asis'}
knitr::kable(storms)
```


## The dplyr::select function
- select: Select/rename variables by name


## The dplyr::select function
- select: Select/rename variables by name
- <b class="myStyle">select(.data, ..., .dots)</b> 

## The dplyr::select function
- select: Select/rename variables by name
- <b class="myStyle">select(.data, ..., .dots)</b> 
    - .data: a tbl

## The dplyr::select function
- select: Select/rename variables by name
- <b class="myStyle">select(.data, ..., .dots)</b> 
    - .data: a tbl
    - ...: comma separated list of unquoted expressions

## The dplyr::select function
- select: Select/rename variables by name
- <b class="myStyle">select(.data, ..., .dots)</b> 
    - .data: a tbl
    - ...: comma separated list of unquoted expressions
    - .dots: Use select_() to do standard evaluation (using strings)


## The dplyr::select function
```{r eval = FALSE}
flights %>% select(year, month)
```

## The dplyr::select function
```{r eval = FALSE}
flights %>% select(year, month)
```
```{r results = 'asis', echo = FALSE}
knitr::kable((flights %>% select(year, month))[1:3,])
```

## The dplyr::select function
- We often have a fixed vector of string variables we would like to keep in the analysis

## The dplyr::select function
- We often have a fixed vector of string variables we would like to keep in the analysis
- Use select_() with .dots argument!

## The dplyr::select function
- We often have a fixed vector of string variables we would like to keep in the analysis
- Use select_() with .dots argument!
```{r eval = FALSE}
flights %>% select_(.dots = c("year", "month", "dep_time", 
                              "dep_delay", "arr_time", 
                              "arr_delay", "flight"))
```
```{r results = 'asis', echo = FALSE}
knitr::kable((flights %>% select_(.dots = c("year", "month", "dep_time", 
                              "dep_delay", "arr_time", 
                              "arr_delay", "flight")))[1:5,])
```

## The dplyr::select function
- Sometimes we might only want to remove a variable or variables
```{r eval = FALSE}
flights %>% select(-year)
flights %>% select_(.dots = c("-year", "-origin", "-day"))
flights %>% select(-sched_dep_time:-arr_time)
```

## Special functions for dplyr::select
- Several variables in your dataset may have a similar naming convention, i.e. variablet1, variablet2, etc. that we want to extract

## Special functions for dplyr::select
- Several variables in your dataset may have a similar naming convention, i.e. variablet1, variablet2, etc. that we want to extract
- Use ends\_with, starts_with, contains, matches, num_range, one_of....


## Special functions for dplyr::select
- Several variables in your dataset may have a similar naming convention, i.e. variablet1, variablet2, etc. that we want to extract
- Use ends\_with, starts_with, contains, matches, num_range, one_of....
```{r eval = FALSE}
## Select variables ending with time and the year/month
flights %>% select(year, month, ends_with("time"))
```{r results = "asis", echo = FALSE}
knitr::kable((flights %>% select(year, month, ends_with("time")))[1:5,1:5])
```

## Special functions for dplyr::select
```{r eval = FALSE}
## Select variables starting with arr and the year/month
flights %>% select(year, month, starts_with("arr"))
```
```{r results = 'asis', echo = FALSE}
knitr::kable((flights %>% select(year, month, starts_with("arr")))[1:8,])
```

## The dplyr::filter function
- Sometimes we need to include only observations which satisfy some criterion for analysis

## The dplyr::filter function
- Sometimes we need to include only observations which satisfy some criterion for analysis
    - Include only cases

## The dplyr::filter function
- Sometimes we need to include only observations which satisfy some criterion for analysis
    - Include only cases
    - Include only those above certain exposure level

## The dplyr::filter function
- Sometimes we need to include only observations which satisfy some criterion for analysis
    - Include only cases
    - Include only those above certain exposure level
    - Include only those with positive survival times

## The dplyr::filter function
- Sometimes we need to include only observations which satisfy some criterion for analysis
    - Include only cases
    - Include only those above certain exposure level
    - Include only those with positive survival times
- The filter function comes to the rescue!

## The dplyr::filter function
- filter: Return rows with matching conditions


## The dplyr::filter function
- filter: Return rows with matching conditions
- <b class = "myStyle">filter(.data, ...)</b> 

## The dplyr::filter function
- filter: Return rows with matching conditions
- <b class = "myStyle">filter(.data, ...)</b>
    - .data: a tbl

## The dplyr::filter function
- filter: Return rows with matching conditions
- <b class = "myStyle">filter(.data, ...)</b>
    - .data: a tbl
    - ...: logical predicates. Multiple conditions are conmbined with &

## The dplyr::filter function
- filter: Return rows with matching conditions
- <b class = "myStyle">filter(.data, ...)</b> 
    - .data: a tbl
    - ...: logical predicates. Multiple conditions are conmbined with &
```{r eval = FALSE}
storms %>% filter(pressure == 1010)
```
```{r results = 'asis', echo = FALSE}
knitr::kable(storms %>% filter(pressure == 1010))
```

## The dplyr::filter function
```{r eval = FALSE}
storms %>% filter(storm == 'Alberto' | pressure > 1009)
```
```{r results = 'asis', echo = FALSE}
knitr::kable(storms %>% filter(storm == 'Alberto' | pressure > 1009))
```

## The dplyr::filter function
- Can use all comparisons, logical expressions in filter(), filter_(). The filter_() function takes string arguments rather than explicit expressions

## The dplyr::filter function
- Can use all comparisons, logical expressions in filter(), filter_(). The filter_() function takes string arguments rather than explicit expressions
- Logical expressions: &, |, !, any, all, ...

## The dplyr::filter function
- Can use all comparisons, logical expressions in filter(), filter_(). The filter_() function takes string arguments rather than explicit expressions
- Logical expressions: &, |, !, any, all, ...
- Comparisons: <, >, ==, <=, >=, !=, %in%, is.na, !is.na, ...

## The dplyr::filter function
- Can use all comparisons, logical expressions in filter(), filter_(). The filter_() function takes string arguments rather than explicit expressions
- Logical expressions: &, |, !, any, all, ...
- Comparisons: <, >, ==, <=, >=, !=, %in%, is.na, !is.na, ...
```{r eval = FALSE}
flights %>% filter(!is.na(air_time))
```
```{r results = 'asis', echo = FALSE}
knitr::kable((flights %>% filter(!is.na(air_time)))[1:4, 1:5])
```

## The dplyr::mutate function
- mutate: Add new variables

## The dplyr::mutate function
- mutate: Add new variables
- <b class = "myStyle">mutate(.data, ...)</b>
    - .data: a tbl
    - ... name-value pairs of expressions, i.e. mean = mean(var)


## The dplyr::mutate function
- mutate: Add new variables
- <b class = "myStyle">mutate(.data, ...)</b>
    - .data: a tbl
    - ...: name-value pairs of expressions, i.e. mean = mean(var)
```{r include = FALSE}
storms %>% mutate(strength = pressure + wind)
```
```{r results = "asis", echo = FALSE}
knitr::kable(storms)
```



## The dplyr::summarise function
- summarise: Summarise multiple values to a single value.

## The dplyr::summarise function
- summarise: Summarise multiple values to a single value.
- <b class = "myStyle">summarise(.data, ...)</b> 

## The dplyr::summarise function
- summarise: Summarise multiple values to a single value.
- <b class = "myStyle">summarise(.data, ...)</b> 
    - .data: a tbl

## The dplyr::summarise function
- summarise: Summarise multiple values to a single value.
- <b class = "myStyle">summarise(.data, ...)</b> 
    - .data: a tbl
    - ...: name-value pairs of summary functions. For example, mean(), sd(), etc.

## The dplyr::summarise function
- summarise: Summarise multiple values to a single value.
- <b class = "myStyle">summarise(.data, ...)</b> 
    - .data: a tbl
    - ...: name-value pairs of summary functions. For example, mean(), sd(), etc.
```{r eval = FALSE}
storms %>% summarise(mean_wind = mean(wind), 
                     mean_pressure = mean(pressure))
```
```{r results = 'asis', echo = FALSE}
knitr::kable(storms %>% summarise(mean_wind = mean(wind), mean_pressure = mean(pressure)))
```

## The dplyr::summarise function
- The dplyr package provides other helpful functions for use with the summarise function

## The dplyr::summarise function
- The dplyr package provides other helpful functions for use with the summarise function
    - first(): first value in a vector 

## The dplyr::summarise function
- The dplyr package provides other helpful functions for use with the summarise function
    - first(): first value in a vector 
    - last(): last value in a vector

## The dplyr::summarise function
- The dplyr package provides other helpful functions for use with the summarise function
    - first(): first value in a vector 
    - last(): last value in a vector
    - nth(): nth value in a vector

## The dplyr::summarise function
- The dplyr package provides other helpful functions for use with the summarise function
    - first(): first value in a vector 
    - last(): last value in a vector
    - nth(): nth value in a vector
    - n(): number of values in a vector 

## The dplyr::summarise function
- The dplyr package provides other helpful functions for use with the summarise function
    - first(): first value in a vector 
    - last(): last value in a vector
    - nth(): nth value in a vector
    - n(): number of values in a vector 
    - n_distinct(): number of distinct values in a vector
    
## The dplyr::summarise function
- The dplyr package provides other helpful functions for use with the summarise function
    - first(): first value in a vector 
    - last(): last value in a vector
    - nth(): nth value in a vector
    - n(): number of values in a vector 
    - n_distinct(): number of distinct values in a vector
- All the usual R functions that take a vector can be used, too!


## The dplyr::group_by function
- Summarizing a data overall is helpful, but often times it is most helpful to provide summaries separately for different categories or values in our data

## The dplyr::group_by function
- Summarizing a data overall is helpful, but often times it is most helpful to provide summaries separately for different categories or values in our data
- <b class = "myStyle">group_by(.data, ..., add = FALSE)</b>

## The dplyr::group_by function
- Summarizing a data overall is helpful, but often times it is most helpful to provide summaries separately for different categories or values in our data
- <b class = "myStyle">group_by(.data, ..., add = FALSE)</b> 
    - .data: a tbl


## The dplyr::group_by function
- Summarizing a data overall is helpful, but often times it is most helpful to provide summaries separately for different categories or values in our data
- <b class = "myStyle">group_by(.data, ..., add = FALSE)</b> 
    - .data: a tbl
    - ...: variables to group by




## The dplyr::group_by function
- Summarizing a data overall is helpful, but often times it is most helpful to provide summaries separately for different categories or values in our data
- <b class = "myStyle">group_by(.data, ..., add = FALSE)</b> 
    - .data: a tbl
    - ...: variables to group by
    - add: If add = FALSE, groups are overridden for tbl .data. Otherwise, groups are added to .data.

## The dplyr::group_by function
```{r eval = F}
flights %>% group_by(dest) %>% filter(!is.na(air_time))
%>% summarise(totalTime = sum(air_time), totalFlights = n())
```

## The dplyr::group_by function
```{r eval = F}
flights %>% group_by(dest) %>% filter(!is.na(air_time))
%>% summarise(totalTime = sum(air_time), totalFlights = n())
```
```{r results = "asis", echo = FALSE}
knitr::kable(flights %>% group_by(dest) %>% filter(!is.na(air_time)) %>% summarise(totalTime = sum(air_time), totalFlights = n()))
```

## The dplyr::group_by function
- Using summarize with group_by makes it really easy to:

## The dplyr::group_by function
- Using summarize with group_by makes it really easy to:
    - retain baseline values via the first() function

## The dplyr::group_by function
- Using summarize with group_by makes it really easy to:
    - retain baseline values via the first() function
    - calculate the number of observations per individual

## The dplyr::group_by function
- Using summarize with group_by makes it really easy to:
    - retain baseline values via the first() function
    - calculate the number of observations per individual
    - calculate the number of distinct observations in an individual for assessing change

## The dplyr::group_by function
- Using summarize with group_by makes it really easy to:
    - retain baseline values via the first() function
    - calculate the number of observations per individual
    - calculate the number of distinct observations in an individual for assessing change
    - Recreate other procedures in R that give statistical summaries by groups (describeBy from the psych package comes to mind...)
- You can always ungroup a tbl by running:
```{r}
## grped_df is grouped
flights %>% ungroup()
## not anymore!
```

## Tidying your data with tidyr        
- The tidyr package is a package created by Hadley Wickham
    
## Tidying your data with tidyr        
- The tidyr package is a package created by Hadley Wickham
    - Defines "tidy" data
    
## Tidying your data with tidyr        
- The tidyr package is a package created by Hadley Wickham
    - Defines "tidy" data
    - Provides rules of thumbs or guidances for creating "tidy" data
    
## Tidying your data with tidyr        
- The tidyr package is a package created by Hadley Wickham
    - Defines "tidy" data
    - Provides rules of thumbs or guidances for creating "tidy" data
    - Read a version of his paper on tidy data by typing ??tidyr in the R console and navigating to the tidyr vignette

## Tidying your data with tidyr        
- The tidyr package is a package created by Hadley Wickham
    - Defines "tidy" data
    - Provides rules of thumbs or guidances for creating "tidy" data
    - Read a version of his paper on tidy data by typing ??tidyr in the R console and navigating to the tidyr vignette
- Tidy data is defined in terms of:
    
## Tidying your data with tidyr        
- The tidyr package is a package created by Hadley Wickham
    - Defines "tidy" data
    - Provides rules of thumbs or guidances for creating "tidy" data
    - Read a version of his paper on tidy data by typing ??tidyr in the R console and navigating to the tidyr vignette
- Tidy data is defined in terms of:
    - variables - hold values and represent constructs, such as height, weight, count, gender....

## Tidying your data with tidyr        
- The tidyr package is a package created by Hadley Wickham
    - Defines "tidy" data
    - Provides rules of thumbs or guidances for creating "tidy" data
    - Read a version of his paper on tidy data by typing ??tidyr in the R console and navigating to the tidyr vignette
- Tidy data is defined in terms of:
    - variables - hold values and represent constructs, such as height, weight, count, gender....
    - observations - one entity representing a unit in our dataset, composed of realizations of one or more variables

## Tidying your data with tidyr

## Tidying your data with tidyr
- Essentially, tidy data satisfies the following requirements

## Tidying your data with tidyr
- Essentially, tidy data satisfies the following requirements
    1. each variable forms a column

## Tidying your data with tidyr
- Essentially, tidy data satisfies the following requirements
    1. each variable forms a column
    2. each observation forms a row

## Tidying your data with tidyr
- Essentially, tidy data satisfies the following requirements
    1. each variable forms a column
    2. each observation forms a row
    3. each type of observational unit forms a table

## Tidying your data with tidyr
- Essentially, tidy data satisfies the following requirements
    1. each variable forms a column
    2. each observation forms a row
    3. each type of observational unit forms a table
- What does "type" of observation refer to?

## Tidying your data with tidyr
- Essentially, tidy data satisfies the following requirements
    1. each variable forms a column
    2. each observation forms a row
    3. each type of observational unit forms a table
- What does "type" of observation refer to?
- Consider an eCommerce customer database. The following tables might exist:

## Tidying your data with tidyr
- Essentially, tidy data satisfies the following requirements
    1. each variable forms a column
    2. each observation forms a row
    3. each type of observational unit forms a table
- What does "type" of observation refer to?
- Consider an eCommerce customer database. The following tables might exist:
    - usernames and passwords

## Tidying your data with tidyr
- Essentially, tidy data satisfies the following requirements
    1. each variable forms a column
    2. each observation forms a row
    3. each type of observational unit forms a table
- What does "type" of observation refer to?
- Consider an eCommerce customer database. The following tables might exist:
    - usernames and passwords
    - purchase records (dates, amounts, products purchased)

## Tidying your data with tidyr
- Essentially, tidy data satisfies the following requirements
    1. each variable forms a column
    2. each observation forms a row
    3. each type of observational unit forms a table
- What does "type" of observation refer to?
- Consider an eCommerce customer database. The following tables might exist:
    - usernames and passwords
    - purchase records (dates, amounts, products purchased)
    - browse records (searches, interests of users)

## Tidying your data with tidyr
- Essentially, tidy data satisfies the following requirements
    1. each variable forms a column
    2. each observation forms a row
    3. each type of observational unit forms a table
- What does "type" of observation refer to?
- Consider an eCommerce customer database. The following tables might exist:
    - usernames and passwords
    - purchase records (dates, amounts, products purchased)
    - browse records (searches, interests of users)
    - demographic information (further information for targeted ads)

## Tidying your data with tidyr
- Essentially, tidy data satisfies the following requirements
    1. each variable forms a column
    2. each observation forms a row
    3. each type of observational unit forms a table
- What does "type" of observation refer to?
- Consider an eCommerce customer database. The following tables might exist:
    - usernames and passwords
    - purchase records (dates, amounts, products purchased)
    - browse records (searches, interests of users)
    - demographic information (further information for targeted ads)
- Tidy data dictates these data be held in separate tables

## Tidying your data with tidyr

## Tidying your data with tidyr
- Why do the extra work of keeping these data separate when I want to analyze them at the same time?

## Tidying your data with tidyr
- Why do the extra work of keeping these data separate when I want to analyze them at the same time?
    - Data changes over time

## Tidying your data with tidyr
- Why do the extra work of keeping these data separate when I want to analyze them at the same time?
    - Data changes over time
    - When you merge/join, your analysis dataset becomes much larger which takes more memory

## Tidying your data with tidyr
- Why do the extra work of keeping these data separate when I want to analyze them at the same time?
    - Data changes over time
    - When you merge/join, your analysis dataset becomes much larger which takes more memory
    - Some information is repeated across many lines, creating more opportunity for error

## The tidyr::gather function

## The tidyr::gather function
- gather: Gather takes multiple columns and collapses into key-value pairs, duplicating all other columns as needed

## The tidyr::gather function
- gather: Gather takes multiple columns and collapses into key-value pairs, duplicating all other columns as needed
- <b class = "myStyle">gather(data, key, value, ...)</b>

## The tidyr::gather function
- gather: Gather takes multiple columns and collapses into key-value pairs, duplicating all other columns as needed
- <b class = "myStyle">gather(data, key, value, ...)</b>
    - data: a data frame (or tbl)

## The tidyr::gather function
- gather: Gather takes multiple columns and collapses into key-value pairs, duplicating all other columns as needed
- <b class = "myStyle">gather(data, key, value, ...)</b>
    - data: a data frame (or tbl)
    - key, value: Names of key and value columns to create in output

## The tidyr::gather function
- gather: Gather takes multiple columns and collapses into key-value pairs, duplicating all other columns as needed
- <b class = "myStyle">gather(data, key, value, ...)</b>
    - data: a data frame (or tbl)
    - key, value: Names of key and value columns to create in output
- Consider dataset cases from the EDAWR package:
```{r results = 'asis', echo = FALSE}
knitr::kable(cases)
```

## The tidyr::gather function

## The tidyr::gather function
- Use the gather function to create a tidy data set

## The tidyr::gather function
- Use the gather function to create a tidy data set
    - Variables are country, year, and count

## The tidyr::gather function
- Use the gather function to create a tidy data set
    - Variables are country, year, and count
    - Each row represents an observation for each country/year pair

## The tidyr::gather function
- Use the gather function to create a tidy data set
    - Variables are country, year, and count
    - Each row represents an observation for each country/year pair
```{r eval = FALSE}
cases %>% gather(year, count, 2:4)
```
```{r results = "asis", echo = FALSE}
knitr::kable((cases %>% gather(year, count, 2:4))[1:5,])
```

## The tidyr::spread function
- Perhaps we have a reason to go the opposite direction

## The tidyr::spread function
- Perhaps we have a reason to go the opposite direction
    - Have one row per observation, but would like to change our unit of observation

## The tidyr::spread function
- Perhaps we have a reason to go the opposite direction
    - Have one row per observation, but would like to change our unit of observation
- spread: Spread a key-value pair across multiple columns

## The tidyr::spread function
- Perhaps we have a reason to go the opposite direction
    - Have one row per observation, but would like to change our unit of observation
- spread: Spread a key-value pair across multiple columns
- <b class = "myStyle">spread(data, key, value, ...)</b>

## The tidyr::spread function
- Perhaps we have a reason to go the opposite direction
    - Have one row per observation, but would like to change our unit of observation
- spread: Spread a key-value pair across multiple columns
- <b class = "myStyle">spread(data, key, value, ...)</b>
    - data: a data frame (or tbl)

## The tidyr::spread function
- Perhaps we have a reason to go the opposite direction
    - Have one row per observation, but would like to change our unit of observation
- spread: Spread a key-value pair across multiple columns
- <b class = "myStyle">spread(data, key, value, ...)</b>
    - data: a data frame (or tbl)
    - key: the bare (unquoted) name of the column whose values will be used as column headings
    
## The tidyr::spread function
- Perhaps we have a reason to go the opposite direction
    - Have one row per observation, but would like to change our unit of observation
- spread: Spread a key-value pair across multiple columns
- <b class = "myStyle">spread(data, key, value, ...)</b>
    - data: a data frame (or tbl)
    - key: the bare (unquoted) name of the column whose values will be used as column headings
    - value: the bare (unquoted) name of the column whose values will populate the cells

## The tidyr::spread function
- Use the spread() function to spread out the year information into separate columns (wide format)

## The tidyr::spread function
- Use the spread() function to spread out the year information into separate columns (wide format)
- Consider the caseslong dataset we just created below:

## The tidyr::spread function
- Use the spread() function to spread out the year information into separate columns (wide format)
- Consider the caseslong dataset we just created below:
```{r echo = FALSE}
caseslong <- cases %>% gather(year,  count, 2:4)
```
```{r results = "asis", echo = FALSE}
knitr::kable(caseslong[1:5,])
```

## The tidyr::spread function
```{r eval = FALSE}
caseslong %>% spread(year, count)
```
```{r results = "asis", echo = FALSE}
knitr::kable((caseslong %>% spread(year, count))[1:5, 1:4])
```

## Summary of R Data Wrangling

## Summary of R Data Wrangling
- Discussion of Data wrangling and it's components

## Summary of R Data Wrangling
- Discussion of Data wrangling and it's components
    - Visualization to come in subsequent lecture using package ggplot2

## Summary of R Data Wrangling
- Discussion of Data wrangling and it's components
    - Visualization to come in subsequent lecture using package ggplot2
- Data manipulation using package dplyr

## Summary of R Data Wrangling
- Discussion of Data wrangling and it's components
    - Visualization to come in subsequent lecture using package ggplot2
- Data manipulation using package dplyr
    - Select variables to trim data set (select())

## Summary of R Data Wrangling
- Discussion of Data wrangling and it's components
    - Visualization to come in subsequent lecture using package ggplot2
- Data manipulation using package dplyr
    - Select variables to trim data set (select())
    - Filter observations to trim data set (filter())

## Summary of R Data Wrangling
- Discussion of Data wrangling and it's components
    - Visualization to come in subsequent lecture using package ggplot2
- Data manipulation using package dplyr
    - Select variables to trim data set (select())
    - Filter observations to trim data set (filter())
    - Mutate dataset by calculating new variables (mutate())

## Summary of R Data Wrangling
- Discussion of Data wrangling and it's components
    - Visualization to come in subsequent lecture using package ggplot2
- Data manipulation using package dplyr
    - Select variables to trim data set (select())
    - Filter observations to trim data set (filter())
    - Mutate dataset by calculating new variables (mutate())
    - Summarizing data set with group_by()

## Summary of R Data Wrangling
- Discussion of Data wrangling and it's components
    - Visualization to come in subsequent lecture using package ggplot2
- Data manipulation using package dplyr
    - Select variables to trim data set (select())
    - Filter observations to trim data set (filter())
    - Mutate dataset by calculating new variables (mutate())
    - Summarizing data set with group_by()
- Data tidying using package tidyr

## Summary of R Data Wrangling
- Discussion of Data wrangling and it's components
    - Visualization to come in subsequent lecture using package ggplot2
- Data manipulation using package dplyr
    - Select variables to trim data set (select())
    - Filter observations to trim data set (filter())
    - Mutate dataset by calculating new variables (mutate())
    - Summarizing data set with group_by()
- Data tidying using package tidyr
    - Convert to long format (gather())
    
## Summary of R Data Wrangling
- Discussion of Data wrangling and it's components
    - Visualization to come in subsequent lecture using package ggplot2
- Data manipulation using package dplyr
    - Select variables to trim data set (select())
    - Filter observations to trim data set (filter())
    - Mutate dataset by calculating new variables (mutate())
    - Summarizing data set with group_by()
- Data tidying using package tidyr
    - Convert to long format (gather())
    - Convert to wide format (spread())



