---
title: "Visualization in R"
author: "Spencer Lourens"
date: '`r format(Sys.Date(), "%B %d %Y")`'
output:
  ioslides_presentation:
    css: ./styles.css
subtitle: The ggplot2 package
fontsize: 10pt
---

```{r library_inlusions_Import_data, include = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(EDAWR)
library(nycflights13)
troops <- read.csv(file = "/Users/spencerlourens/Desktop/Advanced Computing/Intro to ggplot2/troops.csv", header = T)
cities <- read.csv(file = "/Users/spencerlourens/Desktop/Advanced Computing/Intro to ggplot2/cities.csv", header = T)
```

## Outline of Lecture
- Background of ggplot2
- The grammar of graphics
    - Layers
    - Mappings
- Scatterplot Example
    - Geoms
    - Customizing the plot
- Other geom examples
    
## Background of ggplot2
- Data visualiation package and plotting system for R
- Created by Hadley Wickham, Chief Scientist at RStudio
- Takes the "good parts" of lattice and base
    - Can provide plots for many conditions (facet_wrap())
    - Can provide many plots on one screen
- None of the "bad parts"
    - Can add to geoms/annotation plots in "layers", intuitive and simple
    - Overall more intuitive, but sometimes need to look up syntax
- Based on concepts from The Grammar of Graphics by Leland Wilkinson (2005)
- The ggplot2 package has become one of the most widely used in the R language

## Comparing ggplot2 to base graphics / lattice
- Compared to base graphics:
    - ggplot2 is more verbose for simple, canned graphics
    - ggplot2 is less verbose for more complicated, original graphics
- Compared to lattice:
    - ggplot2 provides a more consistent approach across functionalities
    - Lattice has no "grammar of graphics"
    - Both packages can provide "trellis" graphs
        - Display variables or relationships between variables conditioned on another variable (faceting)

## The Grammar of Graphics
- The Grammar of Graphics was published by Leland Wilkinson in 2005
- What do we mean, grammar?
    - "The fundamental principles or rules of an art or science"
    - Essentially, separate visualizations into components, and build visualizations piece-by-piece
- We'll follow Hadley's adaptation of the Grammar of Graphics, employed in ggplot2
    - Data maps to aesthetics which are represented by geoms, creating layers
    - Graphics are composed of one or more layers

## Layers
- Each layer is composed of essentially 4 parts
    - Data and aesthetic mapping
    - A statistical transformation of the data (identity)
    - a geometric object to represent in the graphic
    - a position adjustment (if necessary, i.e. random jitters, dodges)

## Making the Abstract Concrete
- Below is a data frame that we will use to create a scatterplot
```{r echo = FALSE}
age <- rnorm(5, c(12, 14, 16, 17, 19), 0.25)
weight <- rnorm(5, -30 + 10 * age, 0.4)
group <- c("A", "A", "A", "B", "B")
df1 <- data.frame(group = group, age = age, weight = weight)
```
```{r results = "asis", echo = FALSE}
knitr::kable(df1)
```

## Data Defined: now what?
- We may be interested in constructing a scatterplot of weight vs age
- Must define our aesthetics for the plot
```{r echo = FALSE}
df2 <- df1 %>% rename(y = weight, x = age)
```
```{r results = "asis", echo = FALSE}
knitr::kable(df2)
```

## Our First ggplot2 Scatterplot
- Scatterplot code below
```{r eval = FALSE}
ggplot() + geom_point(data = df1, mapping = aes(x = age, 
  y = weight), stat = "identity")
```
```{r echo = FALSE, fig.width = 5, fig.height = 4, fig.align = "center"}
ggplot() + geom_point(data = df1, mapping = aes(x = age, y = weight), 
                      stat = "identity")
```

## Explaining Our First ggplot2 Scatterplot
- The call to ggplot() tells R that we want to use ggplot2 functionality
    - You can specify the dataset here to use the same data for all layers in your plot
    - ggplot(data = df1) + ...
- The call to geom_point() separated by a "+" creates the aesthetic mappings and determines the geom to use
- Argument stat = "identity" tells ggplot to not transform or aggregate the data
    - The stat argument is a little confusing. It doesn't mean, "log", "square root", etc. 
    - Think of this as more of a summary of the data, i.e. "boxplot", "identity", etc.

## Explaining Our First ggplot2 Scatterplot
- The data: df
- The aesthetic mappings: age $\rightarrow$ x, weight $\rightarrow$ y
- The geometric object representing the data: point
- The statistical transformation: identity
- Can add more aesthetic mappings, such as color or facet by group
- Can add more geoms, such as lines, smoothers, etc.

## Our First ggplot2 Scatterplot
```{r eval = FALSE}
ggplot() + 
  geom_point(data = df1, mapping = aes(x = age, y = weight, 
  colour = group), stat = "identity")
```
```{r echo = FALSE, fig.width = 5, fig.height = 4, fig.align = "center"}
ggplot() + geom_point(data = df1, mapping = aes(x = age, y = weight, 
  colour = group), stat = "identity") + ggtitle("Colored by Group")
```

## Our First ggplot2 Scatterplot
```{r eval = FALSE}
ggplot() + 
  geom_point(data = df1, mapping = aes(x = age, y = weight), 
  stat = "identity") + facet_wrap( ~ group)
```
```{r echo = FALSE, fig.width = 5, fig.height = 4, fig.align = "center"}
ggplot() + 
  geom_point(data = df1, mapping = aes(x = age, y = weight), 
  stat = "identity") + facet_wrap( ~ group) + ggtitle("Faceted by Group")
```

## Geometric objects
- Geomertric objects determine how our aesthetic mappings manifest in a visualization
    - Points
    - Bars
    - Lines
    - Paths
    - ...
- The ggplot2 package uses geom_xxx() as a general format for selecting geoms to represent a chosen aesthetic mapping.

## Scales
- Scales control or govern the mapping from data to an aesthetic property
    - continuous, discrete, etc.
- Each aesthetic mapping requires a scale
    - ggplot2 uses defaults scales whenever possible
    - these were employed in our scatterplot example
- Discrete scales are more often necessary to explicitly code for legends
    - scale_colour_discrete(), scale_colour_manual()
    - scale_shape_discrete(), scale_shape_manual()
- Continuous scales can be explicitly coded:
    - scale_x_continuous(), scale_y_continuous

## Our First ggplot2 Scatterplot revisited
```{r eval = FALSE}
ggplot() + 
  geom_point(data = df1, mapping = aes(x = age, y = weight, 
  colour = group), stat = "identity")) + scale_colour_manual(
  values = c("brown", "yellow"))
```
```{r echo = FALSE, fig.width = 5, fig.height = 4, fig.align = "center"}
ggplot() + 
  geom_point(data = df1, mapping = aes(x = age, y = weight, colour = group), 
  stat = "identity") + scale_colour_manual(values = c("brown", "yellow"))
```

## A More Complete Scatterplot
```{r eval = FALSE}
ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width, 
  colour = Species)) + geom_point() + geom_smooth(method = "lm")
```
```{r echo = FALSE, fig.width = 5, fig.height = 4, fig.align = "center"}
ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width, colour = Species)) + geom_point() +   
       geom_smooth(method = "lm")
```

## A More Complete Scatterplot
- A lot more customization possible
    - Remove the confidence bands by setting argument se to FALSE in geom_smooth
    - Change x and y axis names with xlab() and ylab()
    - Add title with ggtitle()
    - Backgrounds can be changed, the grid can be removed, etc.
    - Non-linear smoothers can be used 
    
## A More Complete Scatterplot
```{r eval = FALSE}
ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width, 
  colour = Species)) + geom_point() + geom_smooth(
  method = "loess", span = 0.9) + ylab("Sepal Width") + 
  xlab("Sepal Length") + ggtitle("Sepal Width vs Sepal Length")
```
```{r echo = FALSE, fig.width = 5, fig.height = 3.5, fig.align = "center"}
ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width, colour = Species)) + geom_point() +   
       geom_smooth(method = "loess", span = 0.9) + ylab("Sepal Width") + xlab("Sepal Length") + ggtitle("Sepal Width vs Sepal Length")
```
    
## A Spaghetti Plot in ggplot2
- Using a transformed version of the tb dataset from EDAWR package
```{r echo = FALSE}
tb2 <- tb %>% mutate(cases = adult + child + elderly) %>% group_by(country, year) %>% summarise(cases = sum(cases))
```
```{r eval = FALSE}
ggplot() + geom_line(data = tb2, aes(x = year, 
  y = cases, group = country))
```
```{r warning = FALSE, echo = FALSE, fig.width = 5, fig.height = 3.5, fig.align = "center"}
ggplot() + geom_line(data = tb2, aes(x = year, y = cases, group = country))
```

## Improving our Spaghetti Plot
- Perhaps faceted by country?
```{r eval = FALSE}
ggplot() + geom_line(data = tb2, aes(x = year, 
  y = cases, group = country)) + facet_wrap( ~ country)
```
```{r warning = FALSE, echo = FALSE, fig.width = 5, fig.height = 4, fig.align = "center"}
ggplot() + geom_line(data = tb2[1:304,], aes(x = year, 
  y = cases, group = country)) + facet_wrap(~ country) + theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.text.x = element_text(size = 5))
```

## A Simple Histogram
```{r eval = FALSE, message = FALSE}
ggplot(data = iris, aes(x = Sepal.Width)) + geom_histogram() + 
  geom_point(stat = "bin")
```
```{r warning = FALSE, message = FALSE, echo = FALSE, fig.width = 5, fig.height = 4, fig.align = "center"}
ggplot(data = iris, aes(x = Sepal.Width)) + geom_histogram() + 
  geom_point(stat = "bin")
```

## A Better Looking Histogram 
```{r message = FALSE, eval = FALSE}
ggplot(data = iris, aes(x = Sepal.Width)) + geom_histogram(
  stat = "bin", fill = "grey", colour = "black") + 
  geom_point(stat = "bin", size = 3, colour = "blue")
```
```{r message = FALSE, warning = FALSE, echo = FALSE, fig.width = 5, fig.height = 4, fig.align = "center"}
ggplot(data = iris, aes(x = Sepal.Width)) + geom_histogram(
  stat = "bin", fill = "grey", colour = "black") + 
  geom_point(stat = "bin", size = 3, colour = "blue")
```

## A Better Looking Histogram - Decomposed
- The above plot is composed of two layers
    - A layer using the histogram geom (or rectangle)
    - A layer using the point geom
- Here's the breakdown of each layer
    - data: iris
    - aesthetic: x = Sepal.Width
    - geometric object: histogram/point
    - transformation: "bin" - based on binwidth, bin the values
    - positional adjustment - none

## The Intelligence of ggplot2
- The concept of layering 
- We can "abuse" the intelligence to shorten our code
    - Most geom_xxx() function's have default behavior
    - Don't need to specify stat = "bin" for geom_histogram()
    - Similarly, don't need to specify stat = "boxplot" for geom_boxplot()
    - Don't need stat = "identity" for geom_point(), geom_line()
- If using the same data for each layer, specify data in ggplot() call
- Same for aesthetic mappings

## Recreating a Famous Graphic - Napoleon's March
- Data from Napoleon's historic attempt to defeat Russia
- Often referred to simply as "The French Invasion of Russia"
    - June to December 1812
    - Russia retreated from Smolensk, where Napoleon had a great victory
    - Burned and pillaged their own villages while retreating, preventing Napoleon's army from living off the land
    - Slowly wittled away at his army as they continued to chase the Russian army into the interior of Russia

## Recreating a Famous Graphic - Napoleon's March
- Will recreate Charles Joseph Minard's famous graphic of the march
    - Shows the ever decreasing size of Napoleon's army on their way to Moscow and retreating back to France
    - Minard was a civil engineer who contributed much to information graphics/statistical graphics
    - Small sample of data below
```{r results = "asis", echo = FALSE}
knitr::kable(troops)
```

## The ggplot2 Code 
```{r eval = FALSE}
ggplot(data = troops, aes(x = long, y = lat)) + geom_path(
  aes(size = survivors, colour = direction, group = group), 
  lineend = "round") + geom_text(aes(label = city, y = lat + 0.06), 
  size = 3, data = cities) + geom_point(aes(x = long, y = lat), 
  data = cities) + scale_size(range = c(1,10)) + theme_light() + 
  scale_x_continuous(limits=c(24,39)) + ylab("Latitude") + 
  xlab("Longitude") + scale_colour_manual(values = c("grey50",
  "red")) + ggtitle("Napoleon's March on Moscow")
```

## The Resulting Plot 
```{r echo = FALSE, fig.width = 8.5, fig.height = 5, fig.align = "center"}
ggplot(data = troops, aes(x = long, y = lat)) + geom_path(
  aes(size = survivors, colour = direction, group = group), 
  lineend = "round") + geom_text(aes(label = city, y = lat + 0.06), 
                                 size = 3, data = cities) + 
  geom_point(aes(x = long, y = lat), data = cities) + 
  scale_size(range = c(1,10)) + theme_light() + scale_x_continuous(
  limits=c(24,39)) + ylab("Latitude") + xlab("Longitude") +  scale_colour_manual(values = c("grey50",
  "red")) + ggtitle("Napoleon's March on Moscow")
```
