---
title: "Visualization in R"
author: "Spencer Lourens"
date: '`r format(Sys.Date(), "%B %d %Y")`'
output:
  ioslides_presentation:
    css: ./styles.css
subtitle: The ggplot2 package
fontsize: 10pt
---

```{r library_inlusions, include = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(EDAWR)
library(nycflights13)
```

## Outline of Lecture
- Background of ggplot2
- The grammar of graphics
- Types of plots
- Creating plots
    
    
## Background of ggplot2
- Data visualiation package and plotting system for R
- Created by Hadley Wickham, Chief Scientist at RStudio
- Takes the "good parts" of lattice and base, but none of the "bad parts"
- Based on concepts from The Grammar of Graphics by Leland Wilkinson (2005)
- The ggplot2 package has become one of the most widely used in the R language

## The Grammar of Graphics
- The Grammar of Graphics was published by Leland Wilkinson in 2005
- What do we mean, grammar?
    - "The fundamental principles or rules of an art or science"
    - Essentially, separate visualizations into components, and build visualizations piece-by-piece
- We'll follow Hadley's adaptation of the Grammar of Graphics, employed in ggplot2
    - Data maps to aesthetics to represent geoms, creating layers
    - Graphics are composed of one or more layers

## Layers
- Each layer is composed of essentially 4 parts
    - Data and aesthetic mapping
    - A statistical transformation of the data (identity)
    - a geometric object to represent in the graphic
    - a position adjustment (if necessary, i.e. random jitters, dodges)

## Making the Abstract Concrete
- Below is a data frame that we will use to create a scatterplot
```{r echo = FALSE}
age <- rnorm(5, c(12, 14, 16, 17, 19), 0.25)
weight <- rnorm(5, -30 + 10 * age, 0.4)
class <- c("A", "A", "A", "B", "B")
df1 <- data.frame(group = class, age = age, weight = weight)
```
```{r results = "asis", echo = FALSE}
knitr::kable(df1)
```

## Data Defined: now what?
- We may be interested in constructing a scatterplot of weight vs age
- Must define our aesthetics for the plot
```{r echo = FALSE}
df2 <- df1 %>% rename(y = weight, x = age)
```
```{r results = "asis", echo = FALSE}
knitr::kable(df2)
```

## Our First ggplot2 Scatterplot
- Scatterplot code below
```{r eval = FALSE}
ggplot() + geom_point(data = df1, mapping = aes(x = age, y = weight), 
                      stat = "identity")
```
```{r echo = FALSE, fig.width = 5, fig.height = 4, fig.align = "center"}
ggplot() + geom_point(data = df1, mapping = aes(x = age, y = weight), 
                      stat = "identity")
```

## Explaining Our First ggplot2 Scatterplot
- The data used :df
- The aesthetic mappings: age $\rightarrow$ x, weight $\rightarrow$ y
- The geometric object representing the data: point
- The statistical transformation: identity
- We can add more aesthetic mappings, such as color or facet by class

## Our First ggplot2 Scatterplot
```{r eval = FALSE}
ggplot() + 
  geom_point(data = df1, mapping = aes(x = age, y = weight, 
  colour = class), stat = "identity")
```
```{r echo = FALSE, fig.width = 5, fig.height = 4, fig.align = "center"}
ggplot() + 
  geom_point(data = df1, mapping = aes(x = age, y = weight, 
  colour = class), stat = "identity") + ggtitle("Colored by Group")
```

## Our First ggplot2 Scatterplot
```{r eval = FALSE}
ggplot() + 
  geom_point(data = df1, mapping = aes(x = age, y = weight), 
  stat = "identity") + facet_wrap( ~ group)
```
```{r echo = FALSE, fig.width = 5, fig.height = 4, fig.align = "center"}
ggplot() + 
  geom_point(data = df1, mapping = aes(x = age, y = weight), 
  stat = "identity") + facet_wrap( ~ group) + ggtitle("Faceted by Group")
```

## Geometric objects
- Geomertric objects determine how our aesthetic mappings manifest in a visualization
    - Points
    - Bars
    - Lines
    - Paths
    - ...
- The ggplot2 package uses geom_xxx() as a general format for selecting geoms to represent a chosen aesthetic mapping.

## Scales
- Scales control or govern the mapping from data to an aesthetic property
    - continuous, discrete, etc.
- Each aesthetic mapping requires a scale
    - ggplot2 uses defaults scales whenever possible
    - these were employed in our scatterplot example
- Discrete scales are more often necessary to explicitly code
    - scale_colour_discrete(), scale_colour_manual()
    - scale_shape_discrete(), scale_shape_manual()

## Our First ggplot2 Scatterplot revisited
- We don't like red/blue
```{r eval = FALSE}
ggplot() + 
  geom_point(data = df1, mapping = aes(x = age, y = weight, colour = group), 
  stat = "identity")) + scale_colour_manual(values = c("brown", "yellow"))
```
```{r echo = FALSE, fig.width = 5, fig.height = 4, fig.align = "center"}
ggplot() + 
  geom_point(data = df1, mapping = aes(x = age, y = weight, colour = group), 
  stat = "identity") + scale_colour_manual(values = c("brown", "yellow"))
```
